---
title: Community talk
subtitle: Interactional contexts and consequences
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r include=FALSE}
print_rainbow_flag <- function() {
  if (knitr::is_latex_output()) {
    "rainbow-flag-here"
  } else {
    emo::ji("rainbow_flag")
  }
}
```

Building on everyone's comments and Mako's offer of data, I'll use Reddit, not Twitter, as the starting point for Chapter 3. I really liked the point about thinking about group-level outcomes, though I definitely want to think more about measurable interaction-level outcomes (even if that's just, as someone mentioned, sustained conversation). I'll need to think about the comparative logic of case more -- especially the intersections of place, identity, and activities/interests. I can think of a few example papers using different kinds of subredddits that I might draw on.

## Proposal version

### Short version

Chapter 1 shows that when actual people talk about community they often actually mean something very similar to the sociological concept of Gemeinschaft; chapter 2 explores the conditions under which people experience a sense of community in relation to an identity category. All of this provides evidence of the reality and salience of this sociological thing called "community" for LGBTQ people. But what happens when LGBTQ people seek community out, and express community in the process? I engage with this question in Chapter 3, by using LGBTQ virtual communities to ask who produces community talk, and what happens when they do? 

This chapter will make use of the direct measures of expressions of the social and cultural aspects of community developed in Chapter 1. This will be critical because, as chapter 2 is likely to show, for LGBTQ people, feeling connected to a community is not simply about conformity or sameness. I have chosen to develop my own measure to account for this, rather than reuse existing computational linguistic measures of belonging [@lucy_characterizing_2021], which equate belonging with conformity.

As in Chapter 2, a primary question is whether pro-community factors reinforce each other, or whether community becomes ambient and backgrounded once it has been created. In this chapter, the question is whether community talk is most prevalent among those who are most embedded or central in a group, perhaps performatively creating community for downstream and peripheral group members; or, is it the peripheral members of a group that engage in the most community talk in an effort to create their own sense of belonging? Those are two distinct possibilities for the sources and the potential impacts of expressions of community.

The ideal virtual community for studying community talk would have both network and interaction measures, the contents of conversations, and their eventual outcomes. I don't think perfect data are attainable; instead, different platforms have different downsides. 

Based on initial feedback, I think Twitter comes closest to most of these measures. The problem, of course, is that "gay twitter" doesn't have clear, defined community boundaries. My main idea is to use an identity symbol as a proxy for community membership and a window into the network. (Alternatives like pride hashtags might be worth exploring but seem overly inclusive.)

The fallback plan here is Reddit, which lends itself to a slightly different set of questions; Reddit conveys less information about the embeddedness of individuals in networks of people, but rather in networks of groups. (It has the countervailing advantage that some LGBTQ-oriented groups on Reddit are broader and some are more specialized.)

### Long version

Ultimately, "community" is something that people do -- they build community, they create it. This chapter aims to study how the process of creating and invoking community plays out in social networks and social interactions. Social density of interactions and relations, not mere physical proximity, produces the social reality and the individual subjective experience of community, as well as observable expressions of the same.[^cultural] But then in turn expressions of community help create the conditions for group-based social life. Building on the discursive analysis from chapter 1, and selecting a complementary context compared to the place-based analysis from chapter 2, I aim to use virtual communities as a site to study the relation between social density and expressions of community. However, the contemporary digital platforms that provide spaces for virtual communities are vast and varied, and I need help developing and crystalizing a concrete project that connects digital data sources to some facet of the underlying theoretical process. In this chapter proposal, I'll lay out a few possibilities and articulate their tradeoffs.

[^cultural]: In conjunction with cultural or moral density.

The expression of community I plan to focus on is linguistic or textual, the kind of community talk or vocabulary of belonging that I explored in chapter 1. This could be explicit invocations of community and belonging that keyword methods can surface, or broader language that more generally resonates with those concepts, of the kind that embeddings models can uncover [@stoltz_cultural_2021]. Text is key partly for measurement and analysis reasons; while image, voice, and video data are all also important aspects of the online social experiences that can create or sustain communities, the computational toolkit for using those types of data are less robust and more challenging to apply. In addition to the methodological motivation, there are also important theoretical reasons to focus on text and language, especially in the context of virtual communities that produce informal written language [@mcculloch_because_2019]. There are two ways to think about text: as something that matters in itself, or as a proxy for something else. What comes out of people's mouths (or keyboards) isn't the same thing as what's in their heads. While one view might be that community-oriented language is only interesting as a proxy for the sense of community that individuals feel and perceive internally, in this case I'd argue for the importance of language on its own.This is because I think community talk can be performative, not merely expressive; what's in people's heads can't do anything in the social world unless expressed in some way, and those expressions might have consequences regardless of what individuals really feel.[^both]

[^both]: Of course, I can dream about having both survey measures of what someone perceives they are experiencing internally, and text measures of what they express in social contexts, in order to investigate to what degree those actually correspond, and whether "authentic" expressions operate differently from less authentic ones. I've fielded surveys through social media before, but that kind of measurement seems best left for future work. It would take resources I'm unlikely to have for my dissertation.

A couple analytic considerations shape the possible research questions I might pose. First, do I focus on the causes or consequences of community talk? The structural/interactional formation of a community and the expressions of community that emerge from and in turn reinforce that social reality are a reciprocal, self-reinforcing process that unfolds over time.[^complex] Because of that, disentangling cause and consequence might be difficult. If possible, I want this chapter to focus on the consequences of invoking community, but I'm not committed to that direction yet. That focus would complement the work that chapter 1 and chapter 2 aim to do, by stepping through first what community means, then how community happens, and finally what community does. Still, it might be worth considering preceding contexts and downstream outcomes simultaneously if possible, depending on the exact form the data take. 

[^complex]: i.e., this is a complex system. If need be, simulation might be a way to explore the dynamics better.

Second, there is the matter of scale. There are two temporal scales or levels to consider when thinking about social density. Social density could be structural and relational, adhering in durable, culturally-recognized ties, like a friendship. Or it could be interactional, ephemeral, and activity-based, like a conversation. In a sense, of course, the latter coalesce into and constitute the former, or the former are an emergent, culturally-perceived [@white_notes_2008] property of the latter. The key analytic question is what each time scale lets me observe. It might be easier to observe the outcome of a conversation than to track the evolution of a friendship; it's possible to analyze networks dynamically but more likely I'll get a snapshot of a network than the full temporal dynamic. Interactions often unfold in the context of longstanding ties and/or durable groups, so taking that context into account is warranted regardless of my ultimate focal level of analysis.

With those analytic issues in mind, I'll sketch out a few potential research questions as possible scenarios, in order to make the possibilities inherent in different data sources more concrete. For instance, on the interactional level of a conversation, how might participants bring in or invoke community, whether implicitly or explicitly? People might be chatting back and forth in an amiable way that steadily builds positive emotional energy [@collins_interaction_2004], and that leads to more effusive or expressive community oriented language. Or, community might be invoked in contentious situations (perhaps by a third party), in order to manage contention and steer people toward interactions with more positive emotional energy. In either case, what precedes those moments of heightened community-building? What's the outcome? Backing out further, to the relational level of a group, who are the people who consistently use high amounts of community-oriented language? Are they the most embedded? Accordingly, is combining network and discursive measures a way of identifying people who structurally *and* culturally play a key role in a group [@goldberg_fitting_2016]? Or, are individuals on the margins the most effusive, to compensate for a lack of structural belonging and performatively create their own sense that they belong? I haven't committed yet to any one of those questions or the specific hypotheses they might entail; I expect to be able to refine and align at least one of those questions with some digital data source until they correspond well enough to constitute a feasible and interesting project.

Virtual communities provide a key site for observing how community emerge in interaction. Virtual communities are real sites for building community [@driskell_are_2002; @hampton_neighboring_2003; @rheingold_virtual_2000; @baym_practice_1994], often intertwined with individuals' offline lives as a kind of "augmented reality" [@jurgenson_digital_2011; @orne_boystown_2017] rather than being completely distinct and separate (i.e., "digital dualism"). Moreover, the interactions that constitute virtual community building often happen _through_ text [@mcculloch_because_2019]. By contrast, it's difficult to collect detailed social density data from offline interactions or ties alone [But not impossible, see @boessen_networks_2014]. Another noteworthy aspect of digital spaces is that LGBTQ virtual communities are and have long been highly visible; queer people have been using digital technology to to form connections, to "find" or "build" community with each other, since the virtual communities of the 1990s, like Usenet or the WELL [@dame-griff_herding_2019; @rheingold_virtual_2000; @boyd_its_2014].[^censorship]

[^censorship]: On sexuality and censorship online: one feature of Reddit and Twitter in particular vis-a-vis Facebook, Instagram, TikTok, Tumblr is that they have a more permissive stance toward sexuality (including pornography, but not necessarily other forms of sex work, especially post-SESTA/FOSTA [@blunt_erased_2020]). LGBTQ people frequently come up against the boundaries of censorship of sexuality on the latter set of platforms; they're on all of them, but that's a force pushing them toward the former.

Different digital platforms have distinct structures and affordances, which shape the ways people stage social interactions and the kinds of communities they build. Those affordances also shape the kinds of data I can collect, and ultimately mean I can learn different things from communities on different platforms. In addition, some platforms are better-studied and more convenient to analyze than others; those pragmatic considerations will also shape my case selection and project development. To make the possibilities and tradeoffs clearer, I'll discuss three platforms in detail: Reddit, Wikipedia, and Twitter. In each case, there are subcommunities of LGBTQ people to study, which have unique features but also reveal something about how each platform operates as a whole.

Reddit provides an obvious option for studying virtual communities, because the entire platform is structured around public groups (called _subreddits_). In other words, the key affordance of Reddit is the existence of groups in which to participate; these groups are the focal points for almost all interaction, which occurs in threads of posts and comments within subreddits. Posts and comments can be rated and ranked through upvotes and downvotes, providing a crowdsourced measure of quality [@medvedev_anatomy_2018]. Direct structural ties between people are deemphasized; analyses of networks of Reddit users instead focus on the web of group affiliations [@simmel_georg_1971; @waller_generalists_2019; @waller_quantifying_2021; @olson_navigating_2015] created by ties of subreddit co-membership. In one sense, this group-based structure means that Reddit takes "groupness" for granted -- i.e., it can appear to presume the existence of real community rather than showing how strong or cohesive a given community really is. Everyone who participates to any degree could be said to be a "member" of a "community", which potentially drains those words of any deeper meaning. However, subreddits vary immensely, in size and activity level, but also in moderation efforts, adherence to local rules and norms, and other signals of distinctive group styles and subcultures. Users vary as well, in their levels of participation, how specialized and selective they are in engagement across different groups, and how embedded they are in within-group conversations.

Existing research leverages some of those variations. For instance, @zhang_community_2017 characterize user engagement across a typology of community-level linguistic features; @lucy_characterizing_2021 use contextual word embeddings to study linguistic variation and conformity and identify semantically unique communities; and @waller_quantifying_2021 study polarization of subreddits through "community embeddings" based on co-membership. In the LGBTQ context, Reddit affords opportunities to observe community-oriented language across both large and generic (e.g. r/lgbt) and small and niche (e.g. r/LesbianGamers/) groups. Based on prior work, there are methods to operationalize individual trajectories, conversation "quality" proxies, and so on.

<!-- Footnote: This work is fantastic and often deeply sociological; I don't think sociologists have to cede this terrain just because it touches on digital technology! (My committee gets to decide how much I rant about studying digital things in sociology.) -->

A second, perhaps less intuitively obvious option is Wikipedia. Wikipedia is a public collective representation of knowledge, but Wikipedia is also virtual community; people can come to have a strong sense of identity as Wikipedians [@pentzold_imagining_2011]. This is a "peer production" community, rather than a purely social one. But Wikipedia mobilizes and socializes volunteer labor with strong norms and rules -- to the point of being bureaucratic -- because coordinating the collective production of knowledge is a massive, complicated endeavor (an endeavor the academic community can relate to). On Wikipedia, individual contributors can interact but are not linked by a durable and observable network structure; still, they vary substantially in their overall degree of investment in Wikipedia as a project and community, measurable through time and labor. My primary interest in Wikipedia would be in the backstage interactional work that takes place on the "talk" pages associated with articles, where the outcome of those interactions is the production of public-facing representations. Especially for marginalized identity groups like LGBTQ people, the external stakes of representation are high; moreover, the well-known social biases of Wikipedia in terms of gender, nationality, and so on that emerge from its community of volunteers make clear that this representation is the outcome of a social process [@evans_15_2020].

<!-- TODO: add more citations about Wikipedia, especially from Mako's lab. -->

For instance, English-language Wikipedia has a "List of LGB people." I've previously mapped the latent structure of the articles on this list, using document embeddings. As the t-SNE projection of these embeddings down into two dimensions reveals, these embeddings fractally map substantive domains (e.g., politics, sports, music), nationalities, and genders -- grouping e.g. gay French writers together (Foucault alongside Barthes and Gide and Hocquenghem). Figure 8 shows a thematically labeled version of this map. (An interactive version of the figure which reveals more of its structure is hosted at https://staff.washington.edu/cgilroy/lgb_wikipedia.html). Behind this structure, however, are real people collaborating (and contending) to produce it. They produce different representations with different valences than other Wikipedias in other languages [@park_multilingual_2020]. Other researchers have curated data and produced tools like ConvoKit [@chang_convokit_2020] for analyzing Wikipedia conversations; I might leverage and connect those tools to my own methods for surfacing community language and to the ultimate outcomes of the articles resulting from backstage conversations.

Twitter, the "model organism" of social media [@tufekci_big_2014], is the third major option. While numerous studies based on Twitter data already exist, the platform offers clear advantages for observing community in interaction: both durable friend and follower ties as well as interactions like retweets, quote tweets, and replies are directly observable. In addition, the absence of a clear feature for demarcating the boundaries of groups is in some sense an advantage.[^boundaries] Unlike Reddit, the structure of Twitter doesn't take "community" for granted, making Twitter a more flexible and less predetermined data source. While interactions are structured around around feeds, timelines, and hashtags instead of groups, distinctive Twitter communities and subcultures do exist -- "gay Twitter" among them.

[^boundaries]: Of course, there are distinct conceptual and methodological challenges around boundaries that aren't pregiven. This is one reason many studies of community, social capital, and civil society focus on formal organizations and their membership rather than informal community-building, even though formal organizations are clearly only the tip of the iceberg.

But in order to study how sense of community and social density unfold in some slice of LGBTQ Twitter, I'll need some sort of anchor. One idea I might propose is to use the adoption on Twitter of the rainbow flag, in emoji form (`r print_rainbow_flag()`), either as a window into a broader community, or merely as a way to demarcate a subset of Twitter users of intrinsic interest. While emoji in general can be thought of and analyzed in conjunction with informal written text [e.g., as a form of gesture, @mcculloch_because_2019], `r print_rainbow_flag()` is not just any emoji. Rather, it's a symbol, and a marker of identity. While new symbols proliferate constantly and variations abound, the six-stripe rainbow flag is widely-recognized to the point that it's been formally adopted into the Unicode standard. Twitter users who adopt the `r print_rainbow_flag()` symbol as an identity marker in their bios thus constitute a visible subset of the set of LGBTQ people who use Twitter. And identity is grounds for building community: to the extent that those people belong to a distinct group or community we might call "gay Twitter", I'd expect them to be connected to each other by durable friend/follower ties and interaction-based ties like replies, retweets, and reactions. These social connections should be disproportionately dense, relative to some (as-yet-undetermined) baseline, in a way that network community detection algorithms or clustering metrics would be able to pick out. And from those interactions, sense of community and group culture would be evident and measurable through the same computational text analysis toolkit I've discussed previously.[^geolocated] Compared with the previous potential cases, the outcomes of interactions over time are less clearly defined, but Twitter presents an opportunity to observe a dynamic system as it unfolds.[^time]

<!-- Footnote: Does this include every LGBTQ person on Twitter? Obviously not. Remember, I'm not interested in identity in itself, and this is not a demography project. I don't care to ascribe sexual identities to specific users.  -->

[^geolocated]: Some small fraction of users/tweets will be geolocated as well -- which might present a point of connection to the chapter 2 project about place characteristics.

[^time]: Especially if I'm able to collect data over a long period of time.

However, those three platforms don't exhaust the possibilities for using digital spaces to understand how community operates. In particular, I doubt that they're the contemporary virtual spaces that display the strongest instances of community expression and interactional dynamics. For that, I'd look instead to platforms that facilitate chat or conversation in a fluid, informal way -- a dynamic, rapid temporality lending itself more toward experiences of emotional energy typically associated with co-presence in physical space. This means Discord servers, Slack channels, even Twitch streams or Gather instances or Messenger group chats. Many of those examples are more private, backstage, sometimes ephemeral, and so also offer participants opportunities to express themselves differently; maybe in a less deliberately curated way and in a way that I would think more revealing of interactional dynamics.[^multimodality] Concerns about privacy, ethics, and access limit the feasibility of exploring these alternatives now, but I would not discount them entirely.

[^multimodality]: Some of these newer platforms start to bridge from informal written text to voice and video communication. That multimodality is no doubt an asset for creating a sense of community even as it renders measurement and observation much more complex.

Reddit, Wikipedia, and Twitter present three distinct options for studying social density and community language in an LGBTQ context. These platforms are conventional, well understood, and publicly accessible. This has advantages: the ethical issues and risks of potential additional harm are lessened when many researchers have already studied communities on these platforms, and the power dynamics of being an insider or an outsider to those spaces in varying degrees are perhaps easier to navigate as well. With these platforms, it's more likely I'll be able to construct a comprehensive quantitative picture of a process, and the ways these platforms fail to represent broader populations are also well known. Table 3.1 summarizes the relevant characteristics of Reddit, Wikipedia, and Twitter and how research projects focused on social dynamics and LGBTQ community might be developed from each of them. 

```{r child='table_ch3_proposal.Rmd'}
```
